Supplementary material for "Hack Weeks as a model for Data Science Education and Collaboration"

### 2018-06-21 results.tex
More shorter.

Given the diversity in skills and backgrounds of participants admitted to our hack weeks, a dependence of learning outcomes and teaching on career stage is plausible. At the same time, we suggested earlier in this paper that peer learning is a major mode of knowledge transfer for participants of all career stages. We find no strong evidence for a significant difference between early-career researchers and senior participants for any of the three hack weeks for questions regarding learning outcomes of new tools and topics (Figure \ref{fig:survey}, panel (a)), improvements in day-to-day research (Figure \ref{fig:survey}, (b)) or overall improvements in science (Figure \ref{fig:survey}, (c)) ($p > 0.0007$ (trial-corrected significance threshold) for all with zero or small effect size; data and all (including non-significant) correlations are presented in SM, Section 2, Table 1 and SM Section 3, Figures 1-12). Only for GHW do we find a very weak suggestion that early-career researchers might agree more strongly that the hack week improved their day-to-day research (p=0.02) with a large effect size of $\phi_c = 0.35$ for $1.88$ degrees of freedom (following~\citep{cohen1988}).
Similarly, we find no indication ($p > 0.0007$) that late-career researchers self-report a higher level of teaching at hack weeks compared to early-career researchers (Figure \ref{fig:survey}, (d)). However, the confidence intervals on the measured effect sizes are wide. Testing for the absence of an effect using an equivalence test on the effect size with an equivalence bound corresponding to a moderate effect size suggests that the data are not conclusive enough to reject a moderate to large effect ($ p_\mathrm{eq} > 0.0007$ for all four questions above).

One important question is whether participants from underrepresented groups thrive at hack weeks, or whether their full participation is impeded. Significant differences between minorities and non-minorities, even on a self-reported scale, for questions related to learning outcomes, teaching or network building, would indicate that improvements in workshop facilitation and structure may be required to allow members of these groups to participate fully.
In our surveys, we find no significant dependence of self-reported learning outcomes on gender identity or race/ethnicity ($p > 0.0007$). Similarly, minority participants do not significantly differ in their answers with respect to teaching outcomes for any of the hack weeks ($p > 0.0007$). Attitudes towards building valuable connections at all three hack weeks or the value of their contributions to their hack teams also do not differ between minority and non-minority groups. For GHW, there is a weak indication that participants from racial/ethnic minorities respond more positively when asked about building connections ($p=0.04$) with a medium to large effect size ($\phi_c = 0.4$; $\mathrm{dof}_{\phi_c} = 0.97$), while for AHW, responses regarding the value of contributions to hack teams may be spread more widely for participants from racial/ethnic minorities than for Caucasian participants ($p = 0.01$; $\phi_c = 0.4$, $\mathrm{dof}_{\phi_c} = 0.98$). Equivalence tests reveal systematically small, though non-significant, $p$-values in the range of $p=0.02-0.05$ for AHW for all four questions above in conjunction with gender or ethnic/racial identity, while for GHW and NHW $p > 0.05$ for the same questions (note that the sample size for GHW and NHW was smaller by a factor of 2 compared to AHW). These results, while not a decisive exclusion of an effect of race/ethnicity or gender on hack week participation, provide an initial indication that our careful facilitation strategies may be effective in fostering participation, and that future work on how demographics interact with hack week attendance may be fruitful.

We find that the hack weeks have been largely successful in promoting positive attitudes towards reproducibility and open science: at all three events, the majority reported that the hack week has made them more comfortable with open science  (GHW: 97\%; NHW: 95\%; AHW: 72\%, Figure \ref{fig:survey}, (h)), and more than 85\% of all participants (AHW: 86\%, GHW: 94\%, NHW: 95\%; Figure \ref{fig:survey}, (g)) put code or data created at the hack week into a public repository.
%The overall behaviour hints at how general conventions and attitudes differ in different fields.
%Astronomy shows the largest degree of openness toward open science, whereas our results indicate that open science is still fairly uncommon in the geosciences, with neuroscience falling in between.
%Similar attitudes are reflected when asking whether the hack week has made participants more comfortable with open science (Figure \ref{fig:survey}, (j)): again, geoscience shows the large improvement with over 97\% agreeing with this statement to some degree, followed by neuroscience (95\%) and astronomy (72\%).
%Overall, our results indicate that hack weeks are effective at addressing persisting doubts about making research open and reproducible.
While the focus on open science is not necessarily a required component of a hack week, it aligns naturally with many of the goals and values commonly promoted at hack weeks, such as production of open-source software and data sharing. In some fields, especially where ethical issues around data sharing and privacy are relevant, this might not be a desired focus of the hack week, or might be replaced or augmented by a discussion of ethical considerations.

We caution the reader that our survey results should be read only as an initial indication of the hypotheses we proposed about the use and outcomes of hack weeks, in line with the surveys' exploratory nature. The number of respondents is small and the effects likely subtle. This means that the lack of significant differences may be caused by the small statistical power of our sample. The most important independent variable--attendance at the hack week--is something that our current survey design cannot account for. Moreover, self-reported learning outcomes are not an objective measure, because they are likely subject to internal biases of respondents. Future work will include a more refined survey design as well as attendance versus non-attendance as a control variable.

Because all three events are relatively recent, it is still early to evaluate long-term outcomes, as well as others including publications and collaborations resulting from these events.
There are, however, initial indicators that all hack weeks encouraged long-term engagement with new concepts or tools and that they directly resulted in a number of publications \cite{gullysantiago2015,faria2016,keshavan2017,leonard2017,jordan2017,peterson2017,hahn2017,pricewhelan2017}. For specific examples, see also below.

>>>>

Given the diversity in skills and backgrounds of participants admitted to our hack weeks, a dependence of learning outcomes and teaching on career stage is plausible. At the same time, as suggested earlier, peer learning is a major mode of knowledge transfer for participants of all career stages. We find no strong evidence for a significant difference between early-career and senior participants in any of the hack weeks for questions regarding learning outcomes of new tools and topics (Figure \ref{fig:survey}, panel (a)), improvements in day-to-day research (Figure \ref{fig:survey}, (b)) or overall improvements in science (Figure \ref{fig:survey}, (c)) ($p > 0.0007$ (trial-corrected significance threshold) for all with zero or small effect size; data and all (including non-significant) correlations are presented in SM, Section 2, Table 1 and SM Section 3, Figures 1-12). Only for GHW do we find that early-career researchers agree more strongly that the hack week improved their day-to-day research (p=0.02) with a large effect size of $\phi_c = 0.35$ for $1.88$ degrees of freedom (following~\citep{cohen1988}).
Similarly, we find no indication ($p > 0.0007$) that late-career researchers self-report a higher level of teaching at hack weeks compared to early-career researchers (Figure \ref{fig:survey}, (d)). However, the confidence intervals on the measured effect sizes are wide. Testing for the absence of an effect using an equivalence test on the effect size with an equivalence bound corresponding to a moderate effect size suggests that the data are not conclusive enough to reject a moderate to large effect ($ p_\mathrm{eq} > 0.0007$ for all four questions above).

One important question is whether participants from underrepresented groups thrive at hack weeks, or whether their full participation is impeded. Significant differences between minorities and non-minorities, even on a self-reported scale, for questions related to learning outcomes, teaching or network building, would indicate that improvements in workshop facilitation and structure may be required to allow members of these groups to participate fully.
In our surveys, we find no significant dependence of self-reported learning outcomes on gender identity or race/ethnicity ($p > 0.0007$). Similarly, minority participants do not significantly differ in their answers with respect to teaching outcomes for any of the hack weeks ($p > 0.0007$). Attitudes towards building valuable connections at all three hack weeks or the value of their contributions to their hack teams also do not differ between minority and non-minority groups. For GHW, there is an indication that participants from racial/ethnic minorities respond more positively when asked about building connections ($p=0.04$) with a medium to large effect size ($\phi_c = 0.4$; $\mathrm{dof}_{\phi_c} = 0.97$), while for AHW, responses regarding the value of contributions to hack teams may be spread more widely for participants from racial/ethnic minorities than for Caucasian participants ($p = 0.01$; $\phi_c = 0.4$, $\mathrm{dof}_{\phi_c} = 0.98$). Equivalence tests reveal systematically small, though non-significant, $p$-values in the range of $p=0.02-0.05$ for AHW for all four questions above in conjunction with gender or ethnic/racial identity, while for GHW and NHW $p > 0.05$ for the same questions (note that the sample size for GHW and NHW was smaller by a factor of 2 compared to AHW). These results, while not a decisive exclusion of an effect of race/ethnicity or gender on hack week participation, provide an indication that our facilitation strategies may be effective in fostering participation. Future work on how demographics interact with hack week attendance may be fruitful.

We find that the hack weeks have been largely successful in promoting positive attitudes towards reproducibility and open science: at all three events, the majority reported that the hack week has made them more comfortable with open science  (GHW: 97\%; NHW: 95\%; AHW: 72\%, Figure \ref{fig:survey}, (h)), and more than 85\% of all participants (AHW: 86\%, GHW: 94\%, NHW: 95\%; Figure \ref{fig:survey}, (g)) put code or data created at the hack week into a public repository.
%The overall behaviour hints at how general conventions and attitudes differ in different fields.
%Astronomy shows the largest degree of openness toward open science, whereas our results indicate that open science is still fairly uncommon in the geosciences, with neuroscience falling in between.
%Similar attitudes are reflected when asking whether the hack week has made participants more comfortable with open science (Figure \ref{fig:survey}, (j)): again, geoscience shows the large improvement with over 97\% agreeing with this statement to some degree, followed by neuroscience (95\%) and astronomy (72\%).
%Overall, our results indicate that hack weeks are effective at addressing persisting doubts about making research open and reproducible.
While the focus on open science is not necessarily a required component of a hack week, it aligns naturally with many of the goals and values commonly promoted at hack weeks, such as production of open-source software and data sharing. In some fields, especially where ethical issues around data sharing and privacy are relevant, this should be augmented with a discussion of ethical considerations.

In line with the surveys' exploratory nature, these results should be read only as an initial indication of the hypotheses we proposed about the use and outcomes of hack weeks. The number of respondents is small and the effects likely subtle, and lack of significant differences may be due to statistical power in our sample. Furthermore, the most important independent variable -- attendance of a hack week -- is not accounted for in our current design. Moreover, self-reported learning outcomes are not an objective measure, because they are likely subject to response biases. Future work will include more refined survey design and inclusion of control non-attendees.

Because all three events are relatively recent, it is still early to evaluate long-term outcomes, including publications and collaborations resulting from these events.
There are, however, initial indicators that all hack weeks encouraged long-term engagement with new concepts or tools and that they directly resulted in a number of publications \cite{gullysantiago2015,faria2016,keshavan2017,leonard2017,jordan2017,peterson2017,hahn2017,pricewhelan2017}. Specific examples follow below.
